Predicting the Quality of CDR-H3 Antibody Loop Structural Models

### 2023-03-14 paper.tex
Updates from Overleaf

According to the Kabat definition, the CDR-H3 loop is made up of the
residues H95--H102 (using the Kabat\cite{Kabat1992}, Chothia\cite{Al-Lazikani} or Martin\cite{Abhinandan2008} numbering schemes) in the heavy
chain, with a potential insertion site at position H100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long (Figure~\ref{fig:loopdist}).\lilian{Figure missing!}

@@

Using these modelling methods, framework regions can
generally be predicted with great accuracy (better than 1\AA\
\ca~RMSD\cite{Almagro2014}), as one can often find a very similar
structure for the homology modelling process.  However, the CDR loops
are not as easily predicted due to their great diversity. If the
canonical conformation of CDR loops CDR-L1,L2,L3,H1,H2 can be identified, they too
can be modelled rather well, often within 1\AA\ \ca~RMSD. \highlight{for CDR-H3 loops
the The average values are taken from the antibody modelling assessment average} is usually above 3\AA\cite{Almagro2011}. \lilian{Something odd here --- not sure what it's supposed to say!}

@@

After creating the feature dataset, it is pre-processed (cleaning,
scaling, encoding, see methods for details). Structures with a
resolution worse than 4\AA\ were removed.
Instances of antibodies in our non-redundant dataset that matched in loop sequence were
not removed as 3D-Models \lilian{Just checking you mean models not structures? Do the structures also differ significantly, or do you really mean that the error varies significantly?} of some of these structures with the same loop
sequence differ significantly.
\lilian{If I understand correctly, you mean that CDR-H3 could have the same sequence in multiple structures, but the overall sequence was different?}
The few large RMSD ranges may stem from
low resolution. For example, the loop sequence with the largest RMSD range 
has multiple structure files linked to it of varying quality, one of which has a resolution of only 3.00\AA. Residue differences near the loop may also
explain the conformational difference of the loop itself, 
even if the loop sequence does not differ. Some of these structures are
complexed while others are not, which may also affect the loop
structure (manuscript in preparation).

@@

The RMSD values are also transformed into a set of binary values according to a
list of RMSD thresholds (i.e. a 1 is assigned to above and 0 to below a given threshold). This is done so that binary ML-models can be
trained, which will predict the probability e.g.\ that the 3D-model's RMSD
is above 2\AA, 2.2\AA, 2.4\AA, and so on. \highlight{The number of binary
classifiers incorporated into the first layer} \lilian{You need to describe the overall architecture first so this makes sense. I would have a much simpler version of Figure~\ref{fig:flow} which just shows the ML architecture}  has a great effect on
the final ML-model, the general trend being that the more binary
classifiers are used, the better the nominal prediction.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{flowchart_qualiloop.eps}
  \caption {Simplified pipeline for creating the final machine learning model that will predict 3D-model quality by giving its RMSD range. \lilian{Make Black and WHite!}}
  \label{fig:flow}
\end{figure}

>>>>

According to the Kabat definition, the CDR-H3 loop is made up of the
residues H95--H102 (using the Kabat\cite{Kabat1992}, Chothia\cite{Al-Lazikani} or Martin\cite{Abhinandan2008} numbering schemes) in the heavy
chain, with a potential insertion site at position H100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long.

@@

Using these modelling methods, framework regions can
generally be predicted with great accuracy (better than 1\AA\
\ca~RMSD\cite{Almagro2014}), as one can often find a very similar
structure for the homology modelling process.  However, the CDR loops
are not as easily predicted due to their great diversity. If the
canonical conformation of CDR loops CDR-L1,L2,L3,H1,H2 can be identified, they too
can be modelled rather well, often within 1\AA\ \ca~RMSD, for CDR-H3 loops
the average is usually above 3\AA\cite{Almagro2011}. The average values are taken from the second antibody modelling assessment \cite{Almagro2014}.

@@

After creating the feature dataset, it is pre-processed (cleaning,
scaling, encoding, see methods for details). Structures with a
resolution worse than 4\AA\ were removed.
Instances of antibody structures in our non-redundant dataset that matched in loop sequence (but the overall sequence was different) were
not removed as 3D-Models of some of these structures with the same loop
sequence differ significantly. 
The few large RMSD ranges may stem from
low resolution. For example, the loop sequence with the largest RMSD range 
has multiple structure files linked to it of varying quality, one of which has a resolution of only 3.00\AA. Residue differences near the loop may also
explain the conformational difference of the loop itself, 
even if the loop sequence does not differ. Some of these structures are
complexed while others are not, which may also affect the loop
structure (manuscript in preparation).

@@

The RMSD values are also transformed into a set of binary values according to a
list of RMSD thresholds (i.e. a 1 is assigned to above and 0 to below a given threshold). This is done so that binary ML-models can be
trained, which will predict the probability e.g.\ that the 3D-model's RMSD
is above 2\AA, 2.2\AA, 2.4\AA, and so on. Several of these binary classifiers form a first layer, each outputting a prediction on whether the model is above or below the respective threshold. The outcome values are then fed into a second-layer classifier (Figure~\ref{fig:flow}). The number of binary
classifiers incorporated into the first layer has a great effect on
the final ML-model, the general trend being that the more binary
classifiers are used, the better the nominal prediction.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{simple_flow.eps}
  \caption {Simplified pipeline for creating the final machine learning model that will predict 3D-model quality by giving its RMSD range.}
  \label{fig:flow}
\end{figure}

@@
